---
title: "OLS model"
author: "Rafael Almeida"
date: "12/2/2020"
output: html_document
---

``````{r, message=FALSE, warning = FALSE}
library(tidyverse)
library(lubridate)
library(haven)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(readr)
library(kableExtra)
library(jtools)

```

## Loading the data 

``````{r, message=FALSE, warning = FALSE}

clean_df <- read_csv("clean_df.csv", col_types = cols(date = col_date(format = "%Y-%m-%d")))

```


## Doing some wrangling on "cases"
``````{r, message=FALSE, warning = FALSE}
# Creating a 'cases per 1000 ppl' indicator
clean_df <- clean_df %>%
  mutate(cases_per_1000ppl = ((clean_df$cases / clean_df$total_population) * 1000))
head(clean_df)

# There are some observations that report "cases per 1000 ppl" > 1000. For the purposes of our analyses, we will drop these observations. 
clean_df <- clean_df %>%
  filter(cases_per_1000ppl < 1000)

# We dropped 120 observations
```



## Descriptive statistics

Before devising OLS model(s), we decided to explore potential associations derived from visualizations.


#Question 1 
The epidemiology of the COVID-19 respiratory syndrome suggests that more densely populated counties are more prone to spreading the virus given the airborne nature of the infection. Is this consistent with the data?

``````{r, message=FALSE, warning = FALSE, echo = FALSE}


# Creating the scatter plot of Cases / 1000 ppl over County density

clean_df %>% 
   ggplot(aes(x = pop_density_sq_km, y = cases_per_1000ppl )) +
   geom_point(color = "dark blue") +
   xlab("Pop density / sq_km") + 
   ylab("cases / 1000 pop") +
   labs(title = "Covid-19 cases versus Population Densisty in the U.S in 30-Nov-2020",
              caption = "Data source: XXXXXXX." ) +
   scale_y_continuous(limits = c(0, 1000)) +
  
   theme_bw()
```
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "light blue">
DISCUSSION:
Graphically, we cannot see a strong association between Cases / 1000 population and Population density.
</div>


#Question 2
Continuing to explore the relationship between social distancing and cases, we could hypothesize that more people working from home is associated with less cases. Is this consistent with the data?

``````{r, message=FALSE, warning = FALSE, echo = FALSE}
# Creating a 'cases per 1000 ppl' indicator
clean_df <- clean_df %>%
  mutate(homework_per_1000ppl = ((clean_df$total_pop_worked_home / clean_df$total_population) * 1000))
head(clean_df)

# Creating the scatter plot of Cases / 1000 ppl over Homeworkers / 1000 ppl

clean_df %>% 
   ggplot(aes(x = homework_per_1000ppl, y = cases_per_1000ppl )) +
   geom_point(color = "dark blue") +
   xlab("Home workers / 1000 pop") + 
   ylab("Cases / 1000 pop") +
   labs(title = "Covid-19 cases versus Homeworkers Density in the U.S in 30-Nov-2020",
              caption = "Data source: XXXXXXX." ) +
   scale_y_continuous(limits = c(0, 1000)) +
  geom_smooth(method='lm', formula= y~x) +

   theme_bw()
```
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "light blue">
DISCUSSION:
Graphically, it seems that this relationship should be explored with an OLS,controlled for other covariates. The plot suggests that 'cases / 1000 pop' is positively associated with 'home workers / 1000 pop', which is very counter-intuitive.
</div>

#Question 3
Following the 'standard' protocols for COVID-19 prevention, our third graphical analysis explores the association between mask wearing and Covid-19 incidence. 

``````{r, message=FALSE, warning = FALSE, echo = FALSE}

# Creating a 'high mask usage' indicator
clean_df <- clean_df %>%
  mutate(high_mask_usage = (clean_df$ALWAYS + clean_df$FREQUENTLY))
head(clean_df)


# Creating the scatter plot of Cases / 1000 ppl over Homeworkers / 1000 ppl

clean_df %>% 
   ggplot(aes(x = high_mask_usage, y = cases_per_1000ppl )) +
   geom_line(color = "dark blue") +
   xlab("Mask usage (% of pop)") + 
   ylab("Cases / 1000 pop") +
   labs(title = "Covid-19 cases versus High Mask usage in the U.S in 30-Nov-2020",
              caption = "Data source: XXXXXXX." ) +
   scale_y_continuous(limits = c(0, 1000)) +
  geom_smooth(method='lm', formula= y~x) +

   theme_bw()
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "light blue">
DISCUSSION:
We cannot observe a relevant relationship between high mask-usage and cases / 1000 pop.
</div>

# OLS model assessing Questions 1, 2, and 3
Can we explain the variation in Cases/pop by population density, working from home, and mask - wearing?



``````{r, message=FALSE, warning = FALSE, echo = FALSE}

# Creating the lm 
model1 = lm(cases_per_1000ppl~homework_per_1000ppl + pop_density_sq_km + high_mask_usage, data = clean_df )
summ(model1)
```
